---
title: "Statistics_Practise"
date: "2025-04-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### **13.1(a).** ###
```{r message=FALSE, warning=FALSE}
library(tidyverse)

# Build the data manually as long format
df <- tribble(
  ~Part, ~Operator, ~Trial, ~Measurement,
  1, 1, 1, 50, 1, 1, 2, 49, 1, 1, 3, 50,
  2, 1, 1, 52, 2, 1, 2, 52, 2, 1, 3, 51,
  3, 1, 1, 53, 3, 1, 2, 50, 3, 1, 3, 50,
  4, 1, 1, 49, 4, 1, 2, 51, 4, 1, 3, 50,
  5, 1, 1, 48, 5, 1, 2, 49, 5, 1, 3, 48,
  6, 1, 1, 52, 6, 1, 2, 50, 6, 1, 3, 50,
  7, 1, 1, 51, 7, 1, 2, 51, 7, 1, 3, 51,
  8, 1, 1, 52, 8, 1, 2, 50, 8, 1, 3, 49,
  9, 1, 1, 50, 9, 1, 2, 51, 9, 1, 3, 50,
  10, 1, 1, 47, 10, 1, 2, 46, 10, 1, 3, 49,
  1, 2, 1, 50, 1, 2, 2, 48, 1, 2, 3, 51,
  2, 2, 1, 51, 2, 2, 2, 51, 2, 2, 3, 51,
  3, 2, 1, 54, 3, 2, 2, 52, 3, 2, 3, 51,
  4, 2, 1, 48, 4, 2, 2, 50, 4, 2, 3, 51,
  5, 2, 1, 48, 5, 2, 2, 49, 5, 2, 3, 48,
  6, 2, 1, 52, 6, 2, 2, 50, 6, 2, 3, 50,
  7, 2, 1, 51, 7, 2, 2, 50, 7, 2, 3, 50,
  8, 2, 1, 53, 8, 2, 2, 48, 8, 2, 3, 50,
  9, 2, 1, 51, 9, 2, 2, 48, 9, 2, 3, 49,
  10, 2, 1, 46, 10, 2, 2, 47, 10, 2, 3, 48
)

## Then convert to factors
df <- df %>%
  mutate(
    Part = as.factor(Part),
    Operator = as.factor(Operator)
  )
```


```{r message=FALSE, warning=FALSE}
anova_model <- aov(Measurement ~ Part * Operator, data = df)
summary(anova_model)
```

**Analysis of the Measurement System**

The results of the two-way ANOVA with replication (as shown in the table above) indicate that there is a statistically significant difference between parts (p < 0.001), suggesting the measurement system can effectively distinguish between different part characteristics. 

The operator effect is not statistically significant (p = 0.601), which means both operators performed measurements consistently. 
Furthermore, there is no significant interaction between part and operator (p = 0.927), indicating that the operators measured the parts in a similar manner. 

Overall, the measurement system demonstrates good repeatability and reproducibility, with most of the observed variation attributed to actual part-to-part differences rather than measurement error or operator inconsistencies. 

### **13.1 (b).** ###
```{r message=FALSE, warning=FALSE}
# Constants
n <- 2  # Operators
p <- 10 # Parts
r <- 3  # Trials per Part-Operator

# Mean Squares (from the ANOVA output above)
MS_P <- 11.002
MS_O <- 0.417
MS_PO <- 0.602
MS_E <- 1.500

# Variance components
var_repeatability <- MS_E
var_interaction <- (MS_PO - MS_E) / r
var_operator <- (MS_O - MS_PO) / (n * p)
var_part <- (MS_P - MS_PO) / (n * r)

# Combine into table
tibble(
  Component = c("Repeatability (Error)", "Interaction", "Operator", "Part"),
  Variance = round(c(var_repeatability, var_interaction, var_operator, var_part), 4)
)
```
**Variance Components Estimates**

The estimated variance components are 1.5 for repeatability (error), 1.73 for part-to-part variation, and approximately zero for both the interaction (–0.299) and operator (–0.0092) components.

This indicates that most of the variability comes from part-to-part differences and repeatability, while the operator and interaction effects are negligible.

## **13.6.** ##
```{r variance-fixed-operator}

## Problem 13.6: Operator as a Fixed Effect

### Reanalyze with Operator as a fixed factor (mixed model)


# Reuse constants
n <- 2  # number of operators
r <- 3  # trials
MS_P <- 11.002
MS_PO <- 0.602
MS_E <- 1.500

# Variance components (operator = fixed, so we skip σ²_operator)
var_repeatability <- MS_E
var_interaction <- (MS_PO - MS_E) / r
var_part <- (MS_P - MS_PO) / (n * r)

# Truncate negative estimates to zero
var_interaction <- ifelse(var_interaction < 0, 0, var_interaction)

# Present results
tibble(
  Component = c("Repeatability (Error)", "Part:Operator Interaction", "Part-to-Part"),
  Variance = round(c(var_repeatability, var_interaction, var_part), 4)
)
```

**Reanalyzed with Operators as a Fixed Factor**

When treating Operator as a fixed factor, the estimated variance components indicate that most of the variability is due to part-to-part differences (1.7337), with moderate repeatability error (1.5). The part-operator interaction variance was negative and is interpreted as zero, suggesting no meaningful variation in how operators respond to different parts. This confirms that the measurement system is consistent and capable of distinguishing between parts.

## **6 (a).** Histogram ##
```{r hist-publications}
## Question 6: Faculty Publications Sample

### (a) Histogram & Shape

library(tidyverse)

# Reconstruct the sample from frequency data
pub_data <- tribble(
  ~Publications, ~Frequency,
  0, 28,
  1, 4,
  2, 3,
  3, 4,
  4, 4,
  5, 2,
  6, 1,
  7, 0,
  8, 2,
  9, 1,
  10,1
)

# Expand data
pub_sample <- pub_data %>%
  uncount(Frequency)

# Plot Histogram
ggplot(pub_sample, aes(x = Publications)) +
  geom_histogram(binwidth = 1, boundary = 0, color = "black", fill = "maroon") +
  scale_x_continuous(breaks = 0:10) +
  labs(title = "Histogram of Refereed Publications",
       x = "Number of Publications",
       y = "Number of Faculty") +
  theme_minimal()
```

**Shape of the Data**

This histogram is right-skewed, where the mode is less than the mean. Most faculty members have 0 refereed publications. The next abundant publications by faculty members are between 1-4 which means that most members have 0-4 refereed publications.

## **6(b).** ##
```{r message=FALSE, warning=FALSE}
# Mean number of publications per faculty member
n <- nrow(pub_sample)
y_bar <- mean(pub_sample$Publications)
s <- sd(pub_sample$Publications)
se_y_bar <- s / sqrt(n)

tibble(
  Sample_Size = n,
  Mean_Estimate = round(y_bar, 2),
  Std_Deviation = round(s, 3),
  Std_Error = round(se_y_bar, 3)
)
```
**Mean Number of Publications and Standard Error**

As shown in the table above, the sample mean number of publications is 1.78, with a standard deviation of 2.68 and a standard error of 0.379, based on a sample of 50 faculty members.

### **6 (c).** ###
The sampling distribution of the sample mean will be approximately normal due to the **Central Limit Theorem (CLT)**. Although the data is right-skewed, the sample size is 50, which is sufficiently large for the CLT to apply. Thus, the sample mean can be treated as approximately normal.


### **10. Best SRS Design for Precision** ###

Precision in estimating a population mean improves with **larger sample sizes** relative to the population size. 

Given that all populations have the same variance:

1. **An SRS of size 400 from a population of 4000**  

**Option 1** provides the highest precision, because it has the **largest sampling fraction** (400 / 4000 = 0.1), which most effectively reduces variance in the estimate.
